---
title: "Getting Started with taRantula"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The `taRantula` package provides a production-grade framework for high-volume statistical web scraping. It focuses on **persistence**, **parallelism**, and **compliance**.

## Prerequisites: Selenium Grid

Before scraping, you need a running Selenium Grid. Using Docker is the recommended and easisest way to manage browser overhead outside of `R`.

### Docker-Compose Setup

Save the following as `docker-compose.yaml` and run `docker-compose up -d`. This configuration is optimized for the memory demands of modern browsers. Depending on available resources, additional nodes may be added or the number of allowed sessions should be adjusted.

``` yaml
services:
  hub:
    container_name: selenium-hub
    image: selenium/hub:4.34.0
    ports:
      - 4442-4444:4442-4444
    networks:
      - tarantula-network
    environment:
      - SE_NODE_SESSION_TIMEOUT=600
      - SE_SESSION_REQUEST_TIMEOUT=300
      - SE_NODE_CONNECTION_TIMEOUT=60
      - SE_NODE_READ_TIMEOUT=60

  node-chrome-1:
    image: selenium/node-chrome:138.0
    container_name: selenium-grid-chrome-1
    depends_on:
      - hub
    networks:
      - tarantula-network
    environment:
      - SE_EVENT_BUS_HOST=hub
      - SE_EVENT_BUS_PUBLISH_PORT=4442
      - SE_EVENT_BUS_SUBSCRIBE_PORT=4443
      - SE_NODE_OVERRIDE_MAX_SESSIONS=true
      - SE_NODE_MAX_SESSIONS=3
      - SE_ENABLE_BROWSER_LEFTOVERS_CLEANUP=true
      - SE_BROWSER_LEFTOVERS_INTERVAL_SECS=600
      - SE_BROWSER_LEFTOVERS_TEMPFILES_DAYS=2
    shm_size: 2g
    deploy:
      resources:
        limits:
          cpus: '2.5'
          memory: 6G
        reservations:
          cpus: '0.5'
          memory: 2G

networks:
  tarantula-network:
  external: true
```

***Note**:* Create the network first if needed: This can be achieved using command `docker network create tarantula-network`.

## Basic Workflow

The package contains the following processes: 
    
- define search parameters via an `R6`-based configuration and run Google Custom Search or Brave Search.
- define scraping parameters via an `R6`-based configuration and initializing the scraper.
    

## Search URL Candidates

### Configuration 

For Google Custom Search, set the following environmental variables:

  - `SCRAPING_APIKEY_GOOGLE` for the google custom search api key
  - `SCRAPING_ENGINE_GOOGLE` for the google custom search engine ID

For Brave Search, set `SCRAPING_APIKEY_BRAVE` for the Brave Search API key.

Use `paramsGoogleSearch()` or `paramsBraveSearch()` to manage settings. This ensures all nested parameters are validated before the search starts.

### Google Custom Search

Run Google Custom Search API with queries built from enterprise name and addresses.

```{r}
Sys.setenv(
  SCRAPING_APIKEY_GOOGLE = "My_ApiKey",
  SCRAPING_ENGINE_GOOGLE = "My_Engine"
)

cfg <- paramsGoogleSearch(
  path = "~/path/to/my/project",
  scrape_attributes = c("title", "link", "displayLink", "snippet")
)

dat <- data.table(
  ID = c(1, 2, 3),
  EnterpriseName = c("Name1", "Name2", "Name3"),
  EnterpriseAddress = c("Address1", "Address2", "Address3")
)

# build queries
dat[, Query1 := buildQuery(.SD), .SDcols = c(
  "EnterpriseName",
  "EnterpriseAddress"
)]
dat[, Query2 := buildQuery(.SD), .SDcols = c("EnterpriseAddress")]

# update keys in config
cfg$set(key = "query_col", c("Query1", "Query2"))
cfg$set(key = "file", c("File1.csv", "File2.csv")) # <- files to save results in

# run google custom search
runGoogleSearch(
  cfg = cfg,
  dat = dat
)
```

### Brave Search with a URL blacklist

Brave Search supports URL blacklists through the `blacklisted_urls` setting. The domains or URLs are written to a temporary `.goggle` file and sent to the Brave Search API as Goggles rules.

```{r}
Sys.setenv(SCRAPING_APIKEY_BRAVE = "My_Brave_ApiKey")

dat <- data.table(
  id = 1,
  term = "st-georgen-kreischberg"
)
dat[, query := buildQuery(.SD), .SDcols = "term"]

cfg <- paramsBraveSearch(
  id_col = "id",
  query_col = "query",
  blacklisted_urls = "https://www.st-georgen-kreischberg.gv.at/",
  scrape_attributes = c("title", "link", "displayLink", "snippet"),
  verbose = FALSE
)

urls <- searchURL(
  cfg = cfg,
  dat = dat,
  file = NULL,
  query_col = "query"
)
```

## Scraping

### Configuration 

Use `paramsScraper()` to manage settings. This ensures all nested parameters are validated before the scraper starts. The `UrlScraper` class manages the state, connection to the `DuckDB` database, and the parallel worker pool.

```{r setup-config}
library(taRantula)

# Initialize scraping parameters
cfg <- paramsScraper()
cfg$set("selenium$host", "localhost")
cfg$set("selenium$workers", 3)
cfg$set("urls", c("https://nytimes.com", "https://whitehouse.gov"))

# Number of workers for robots.txt detection
cfg$set("robots$workers", 4)

# Initialize the scraper
s <- UrlScraper$new(cfg)

# Start scraping
s$scrape()
```

### Scraping Engines: Selenium vs. httr2

The `scrape()` function supports two engines: Selenium for dynamic content and `httr2` for static web requests. By default, the package uses Selenium. You can switch to `httr2` by setting the configuration parameter `cfg$set("selenium$use_selenium", FALSE)`.

#### Selenium Mode (JavaScript Support)
By default, the scraper expects a **Selenium Grid** to work with. This is necessary for modern websites that require JavaScript to render content.

* **Redirect Detection**: The scraper automatically detects if a URL redirects (e.g., from `http` to `https`) and stores the `url_redirect` pointer in the database.

* **Content Persistence**: The full rendered HTML source is captured after the DOM has stabilized.

#### httr2 Fallback (High Speed)
If you disable Selenium, the scraper falls back to `httr2::request()`. This is significantly faster and uses fewer resources but does not allow to catch dynamically rendered content.

* **When to use**: For static sites or APIs where JavaScript rendering is not required.
* **Custom Headers**: In this mode, the scraper passes your configured user-agent and metadata via HTTP headers.

```{r engine-toggle}
# To disable Selenium and use the httr2 fallback:
cfg$set("selenium$use_selenium", FALSE)

# The same UrlScraper$new(cfg) will now use httr2 internally
```

### Fault Tolerance and Progress Protection

One of the core strengths of `taRantula` is its resilience. Scraping thousands of URLs is prone to interruptions, such as network timeouts, hardware crashes, or OS updates.

#### Snapshotting Mechanism

The configuration includes a `snapshot_every` parameter (default is 10).

**How it works:** Every N URLs, each worker writes its current progress to a temporary "snapshot" file in the project directory. Simultaneously, all extracted data is persisted into highly compressed **Parquet files**, ensuring the actual scraped content is saved immediately.

**Crash Recovery:** If the `R` session or the Selenium server crashes, simply re-initialize the `UrlScraper` with the same configuration. On startup, the class detects these snapshot files and reconciles them with the existing DuckDB database and Parquet archives.

**Persistence:** The snapshotting mechanism ensures that only the URLs currently in the buffer are at risk; since all successfully scraped data is written to Parquet files during the operation, a large percentage of the progress is permanently preserved.

### Data Extraction and Iteration

Data is managed through a hybrid architecture: relational metadata (status, timestamps, and hierarchy) is stored in **DuckDB**, while the raw extracted content (HTML source and discovered link lists) is archived in partitioned **Parquet files**.

The `$results()` method abstracts this complexity by automatically performing an on-the-fly join between the DuckDB metadata and the corresponding Parquet archives. For large datasets, you can toggle the `with_src` parameter to choose between high-performance metadata analysis and full-content extraction.

```{r extraction}
# By default: Access results as a data.table, including the HTML source code.
# Note: This performs an on-the-fly join with the Parquet archive.
results_dt <- s$results()

# Access metadata only. 
# This is significantly faster for filtering or summary statistics.
results_meta <- s$results(with_src = FALSE)

# Access all discovered links; This internally computes 
# hierarchy levels
links_dt <- s$links()

# Inspect execution logs for progress tracking and redirects
logs_dt <- s$logs()
```

#### Advanced Extraction: Regex

You can also perform targeted data mining across your results. For example, to find an Austrian UID (VAT number) only on pages that are likely to be "Imprints":

```{r regex-example}
# Pattern for Austrian VAT numbers
uid_pattern <- "ATU[0-9]{8}"

# Extract only from pages where the URL or the link label matches 'imprint'
vat_results <- s$regex_extract(
  pattern = uid_pattern,
  filter_links = "imprint",
  ignore_cases = TRUE
)
```

### Controlling the Scraper

#### Adding New URLs

You can feed newly discovered links back into the scraper. `taRantula` automatically handles duplicate detection.

```{r update}
# Add new URLs and scrape again
s$update_urls(urls = c("https://google.com", "https://www.statistik.at", "https://ec.europa.eu/eurostat"))
s$scrape()
```

#### Removing Pending URLs

If you accidentally queued links or want to prune the queue before the next scraping run, you can remove URLs that are still marked as `"todo"`. Any links that have already been scraped or blocked will remain untouched.

```{r remove}
# Remove specific pending URLs from the queue
s$remove_urls(urls = c("https://ec.europa.eu/eurostat"))
```

#### Graceful Termination

If you need to stop a long-running scrape, use `$stop()`. This creates a signal file that tells workers to finish their current URL and exit cleanly.

```{r stop}
s$stop()
```

#### Cleanup

Always close the scraper to shut down the `DuckDB` connection and clean up temporary files.

```{r close}
s$close()
```

## Resources

-   [Google Custom Search API](https://developers.google.com/custom-search/v1/overview)
-   [Docker Engine Installation](https://docs.docker.com/engine/install/)
-   [Selenium Grid Documentation](https://www.selenium.dev/documentation/grid/)
